---
title: "Appendix"
format:
  pdf:
    include-in-header:
      text: |
        \usepackage{changepage}
        \usepackage[top=1cm, bottom=1cm, left=1.5cm, right=2.5cm]{geometry}
editor: visual
---

\Large

All relevant files can be found by visiting the following link:

[github.com/shatakshishelar/TippecanoeCounty](github.com/shatakshishelar/TippecanoeCounty)

That link contains:

-   All 2023 ratio study results (Countywide & Clustered)

-   Some 2024 ratio study results (Countywide from available counties)

-   All plots displayed in the Appendix

-   Combined Workbooks, Ratio Study Files, and Sales Reconciliation files used in our analysis

-   R code files used to perform analysis and generate relevant plots

    \newpage

# 1. Percentage of Sales Excluded

\large

This table shows the percentages of available sales that each county excluded on the 2023 ratio studies.

```{r}
#| echo: false
#| message: false
#| warning: false
#| results: asis
library(knitr)
library(dplyr)
library(stringr)
library(tidyr)
exc23 = read.csv("tippecanoe/Combined_Reconciliation_2023.csv")
exc24 = read.csv("tippecanoe/Combined_Reconciliation_2024.csv")
rat23 = read.csv("tippecanoe/Combined_Study_2023.csv")
rat24 = read.csv("tippecanoe/Combined_Study_2024.csv")
exc23 = rbind(exc23,exc24[exc24$County == 'Whitley',])
rat23 = rbind(rat23,rat24[rat24$County == 'Whitley',])
exc23 = exc23[!is.na(exc23$ParcelNumber),]
rat23 = rat23[!is.na(rat23$ParcelNumber),]

# Step 1: Count excluded sales by county
excluded_counts <- exc23 %>%
  group_by(County) %>%
  summarise(Excluded = n(), .groups = 'drop')

# Step 2: Count retained (study) sales by county
retained_counts <- rat23 %>%
  group_by(County) %>%
  summarise(Retained = n(), .groups = 'drop')

# Step 3: Join and calculate percentage excluded
sales_summary <- full_join(excluded_counts, retained_counts, by = "County") %>%
  mutate(
    TotalExcluded = replace_na(Excluded, 0),
    Retained = replace_na(Retained, 0),
    Total = TotalExcluded + Retained,
    PercentExcluded = round(100 * TotalExcluded / Total, 1)
  ) %>%
  select(County, PercentExcluded,TotalExcluded) %>%
  arrange(desc(PercentExcluded))

latex_table1 <- kable(sales_summary[1:44, ], format = "latex", booktabs = TRUE,row.names = FALSE)
cat("\\begingroup\n\\begin{center}\n\\setlength{\\tabcolsep}{3pt}\n\\normalsize\n", latex_table1, "\n\\end{center}\n\\endgroup\n")

latex_table2 <- kable(sales_summary[47:92, ], format = "latex", booktabs = TRUE,row.names = FALSE)
cat("\\begingroup\n\\begin{center}\n\\setlength{\\tabcolsep}{3pt}\n\\normalsize\n", latex_table2, "\n\\end{center}\n\\endgroup\n")

```

\Large

# 2. Ratio Study Summaries

These tables show the countywide results by each ratio study category. County by county results are shown in section 2. These tables facilitate comparisons among counties, and reveal statewide trends.

```{r}
#| echo: false
#| message: false
#| warning: false
#| results: asis
library(knitr)
library(dplyr)
library(stringr)

county_data <- read.csv("tippecanoe/All_Counties_2023.csv")
ogboundfind = read.csv("Countywide_Original_2023.csv")
modboundfind = read.csv("Countywide_Moderate_2023.csv")
agboundfind = read.csv("Countywide_Aggressive_2023.csv")
ogboundfind = ogboundfind[ogboundfind$Grouping == 'Countywide',c(5,7)]
modboundfind = modboundfind[modboundfind$Grouping == 'Countywide',c(5,7)]
agboundfind = agboundfind[agboundfind$Grouping == 'Countywide',c(5,7)]
library(dplyr)

ogboundfind <- ogboundfind %>%
  mutate(
    `Org Bounds` = sprintf("%.2f - %.2f", round(LowerBound, 2), round(UpperBound, 2))
  )
modboundfind <- modboundfind %>%
  mutate(
    `Mod Bounds` = sprintf("%.2f - %.2f", round(LowerBound, 2), round(UpperBound, 2))
  )
agboundfind <- agboundfind %>%
  mutate(
    `Agr Bounds` = sprintf("%.2f - %.2f", round(LowerBound, 2), round(UpperBound, 2))
  )
county_data = cbind(county_data,ogboundfind,modboundfind,agboundfind)
for (type in c('Industrial Vacant','Industrial Improved','Commercial Vacant','Commercial Improved','Residential Vacant','Residential Improved')) {
  county_sub = county_data[county_data$Type == type,]
  county_sub = county_sub[county_sub$`Agr.Sales` > 4,]
  county_data1 <- county_sub[, c(2,4,5,6,18,21,24,7,8,9,13,14,15)]
  latex_table1 <- kable(county_data1, format = "latex", booktabs = TRUE,row.names = FALSE)
  cat("\\centering\n",type,"")
  # Wrap in LaTeX font size environment
  cat("\\begin{adjustwidth}{-0.5cm}{0cm}\n")
  cat("\\begingroup\n\\setlength{\\tabcolsep}{3pt}\n\\tiny\n", latex_table1, "\n\\endgroup\n")
  cat("\\end{adjustwidth}\n")



}
```

```{=tex}
\begin{flushleft}
```
\Large

# 3. County Summaries

For each county in Indiana, you will find a summary page complete with:

-   a county map showing the rural (red) and urban (blue) defined areas based on our zip code classification system using parcel density,

-   a list of clusters that the county appears in,

-   the median, COD, PRD, and number of sales for each of the six ratio study categories based on our three unique calculations,

-   and agricultural statistics for 2022, 2023, and 2024.

Note: White areas may appear on the map of some counties. Some of these areas represent exempt parcels (Government land, State Parks, etc.) while others may have been lost as we joined together the workbook and shape files for each county. These areas were treated as though they did not exist for the purpose of rural and urban classifications. Anomalies (rural parcels in urban zip codes or urban parcels in rural zip codes) also occur in some counties, likely due to incorrectly recorded zip codes within one or more of the files used. These anomalies were ruled to be too inconsequential to manually hunt down, so they were left as they are.

```{=tex}
\end{flushleft}
```
\newpage

```{r}
#| echo: false
#| message: false
#| warning: false
#| results: asis
library(knitr)
library(dplyr)
library(stringr)

county_data <- read.csv("tippecanoe/All_Counties_2023.csv")
clustersource = read.csv("Clustered_Aggressive_2023.csv")
ogboundfind = read.csv("Countywide_Original_2023.csv")
modboundfind = read.csv("Countywide_Moderate_2023.csv")
agboundfind = read.csv("Countywide_Aggressive_2023.csv")
ogboundfind = ogboundfind[ogboundfind$Grouping == 'Countywide',c(5,7)]
modboundfind = modboundfind[modboundfind$Grouping == 'Countywide',c(5,7)]
agboundfind = agboundfind[agboundfind$Grouping == 'Countywide',c(5,7)]
library(dplyr)

ogboundfind <- ogboundfind %>%
  mutate(
    `Org Bounds` = sprintf("%.2f - %.2f", round(LowerBound, 2), round(UpperBound, 2))
  )
modboundfind <- modboundfind %>%
  mutate(
    `Mod Bounds` = sprintf("%.2f - %.2f", round(LowerBound, 2), round(UpperBound, 2))
  )
agboundfind <- agboundfind %>%
  mutate(
    `Agr Bounds` = sprintf("%.2f - %.2f", round(LowerBound, 2), round(UpperBound, 2))
  )
county_data = cbind(county_data,ogboundfind,modboundfind,agboundfind)

agdata = read.csv("Countywide_Agricultural_Study.csv",row.names=NULL)
filepath <- "tippecanoe/county_pictures/zip_classification_"
get_groupings_for_county <- function(county_name) {
  groupings <- clustersource %>%
    filter(str_detect(County, paste0("\\b", county_name, "\\b"))) %>%
    pull(Grouping) %>%
    unique() %>%
    sort()

  cat("\\centering\nClusters:", paste(groupings, collapse = ", "), "\n")
}

for(county in unique(county_data$County)) {
  image_path <- paste0(filepath, county, ".jpg")
  cat("![](", image_path, ")\n")
  county_sub = county_data[county_data$County == county,]
  county_data1 <- county_sub[, c(3,4,5,6,18,21,24)]
  county_data2 = county_sub[, c(3,7,8,9,10,11,12,13,14,15)]
  latex_table1 <- kable(county_data1, format = "latex", booktabs = TRUE,row.names = FALSE)
  cat("Countywide Data:")
  # Wrap in LaTeX font size environment
  cat("\\begingroup\n\\setlength{\\tabcolsep}{3pt}\n\\small\n", latex_table1, "\n\\endgroup\n")


  latex_table2 <- kable(county_data2, format = "latex", booktabs = TRUE,row.names = FALSE)
  
  # Wrap in LaTeX font size environment
  cat("\\begingroup\n\\setlength{\\tabcolsep}{3pt}\n\\small\n", latex_table2, "\n\\endgroup\n")

  
  # Example: Output for "Allen"
  get_groupings_for_county(county)
  
  rownames(agdata) <- NULL
  agdata_sub = agdata[agdata$County == county,]
  agdata_sub = agdata_sub[,c(3,4,5,6,7,8,10,12)]
  agdata_sub <- agdata_sub[, colnames(agdata_sub) != ""]
  latex_table3 <- kable(agdata_sub, format = "latex", booktabs = TRUE,row.names = FALSE)
  cat("\nAgricultural Data:\n")
  cat("\\begingroup\n\\setlength{\\tabcolsep}{3pt}\n\\small\n\\centering\n")
  cat(latex_table3, "\n\\endgroup\n")



  
}
```

\newpage

# 4. Cluster Summaries

```{=tex}
\begin{flushleft}

For each cluster, you will find a summary page complete with:

\end{flushleft}
```
-   a map showing the highlighted cluster,

-   a list of counties that appear in the cluster,

-   the median, COD, PRD, and number of sales for each of the six property types based on our three unique calculations,

-   and confidence interval box plots for each of the six ratio study categories, as well as each property class comprising that category, based on our three unique calculations.

    \newpage

```{r}
#| echo: false
#| message: false
#| warning: false
#| results: asis

cluster_data <- read.csv("tippecanoe/All_Clusters_2023.csv")
clustersource = read.csv("Clustered_Aggressive_2023.csv")
ogboundfind = read.csv("Clustered_Original_2023.csv")
modboundfind = read.csv("Clustered_Moderate_2023.csv")
agboundfind = read.csv("Clustered_Aggressive_2023.csv")
ogboundfind = ogboundfind[order(ogboundfind$Grouping),]
modboundfind = modboundfind[order(modboundfind$Grouping),]
agboundfind = agboundfind[order(agboundfind$Grouping),]
ogboundfind = ogboundfind[,c(5,7)]
modboundfind = modboundfind[,c(5,7)]
agboundfind = agboundfind[,c(5,7)]


ogboundfind <- ogboundfind %>%
  mutate(
    `Org Bounds` = sprintf("%.2f - %.2f", round(LowerBound, 2), round(UpperBound, 2))
  )
modboundfind <- modboundfind %>%
  mutate(
    `Mod Bounds` = sprintf("%.2f - %.2f", round(LowerBound, 2), round(UpperBound, 2))
  )
agboundfind <- agboundfind %>%
  mutate(
    `Agr Bounds` = sprintf("%.2f - %.2f", round(LowerBound, 2), round(UpperBound, 2))
  )
cluster_data = cbind(cluster_data,ogboundfind,modboundfind,agboundfind)

picfilepath <- "tippecanoe/cluster_pictures/cluster_"
get_counties_for_cluster <- function(cluster_number) {
  counties <- clustersource %>%
    filter(Grouping == cluster_number) %>%
    pull(County) %>%
    str_split(",\\s*") %>%     # split comma-separated county strings
    unlist() %>%
    str_trim() %>%
    .[. != ""] %>%
    unique() %>%
    sort()

  cat("Counties:", paste(counties, collapse = ", "), "\n")
}


for(cluster in unique(cluster_data$Cluster)) {
  image_path <- paste0(picfilepath, cluster, ".jpg")
  cat("![](", image_path, ")\n")
  get_counties_for_cluster(cluster)
  cluster_sub = cluster_data[cluster_data$Cluster == cluster,]
  cluster_data1 <- cluster_sub[, c(3,4,5,6,18,21,24)]
  cluster_data2 = cluster_sub[, c(3,7,8,9,10,11,12,13,14,15)]
  latex_table1 <- kable(cluster_data1, format = "latex", booktabs = TRUE,row.names = FALSE)
  cat("\nCluster Data:\n")
  # Wrap in LaTeX font size environment
  cat("\\begingroup\n\\setlength{\\tabcolsep}{3pt}\n\\small\n", latex_table1, "\n\\endgroup\n")

  latex_table2 <- kable(cluster_data2, format = "latex", booktabs = TRUE,row.names = FALSE)
  
  # Wrap in LaTeX font size environment
  cat("\\begingroup\n\\setlength{\\tabcolsep}{3pt}\n\\small\n", latex_table2, "\n\\endgroup\n")

  cat("\\newpage\n")

  # Header for next section
  cat("Original Boxplots:\n")
  image_path <- paste0("tippecanoe/original_boxplots/", cluster, "_Overall.jpg")
  cat("\n![](", image_path, ")\n")
  image_path <- paste0("tippecanoe/original_boxplots/", cluster, "__1.jpg")
  cat("![](", image_path, ")\n")
  image_path <- paste0("tippecanoe/original_boxplots/", cluster, "__2.jpg")
  cat("![](", image_path, ")\n")
  
  cat("\\newpage\n")

  # Header for next section
  cat("Moderate Boxplots:\n")
  image_path <- paste0("tippecanoe/moderate_boxplots/", cluster, "_Overall.jpg")
  cat("\n![](", image_path, ")\n")
  image_path <- paste0("tippecanoe/moderate_boxplots/", cluster, "__1.jpg")
  cat("![](", image_path, ")\n")
  image_path <- paste0("tippecanoe/moderate_boxplots/", cluster, "__2.jpg")
  cat("![](", image_path, ")\n")
  
  cat("\\newpage\n")

  # Header for next section
  cat("Aggressive Boxplots:\n")
  image_path <- paste0("tippecanoe/aggressive_boxplots/", cluster, "_Overall.jpg")
  cat("\n![](", image_path, ")\n")
  image_path <- paste0("tippecanoe/aggressive_boxplots/", cluster, "__1.jpg")
  cat("![](", image_path, ")\n")
  image_path <- paste0("tippecanoe/aggressive_boxplots/", cluster, "__2.jpg")
  cat("![](", image_path, ")\n")

}
```

\newpage

# References

Indiana Department of Local Government Finance. (2024, December 2). Ratio Study. <https://www.in.gov/dlgf/assessments/ratio-study/.>

International Association of Assessing Officers. Standard on Ratio Studies. International Association of Assessing Officers, Apr. 2013, [www.iaao.org/wp-content/uploads/Standard_on_Ratio_Studies.pdf](www.iaao.org/wp-content/uploads/Standard_on_Ratio_Studies.pdf)

Kuethe, T. H., Langemeier, M., & Mintert, J. (2024, August). 2024 Purdue Farmland Values and Cash Rents Survey Results. Purdue University Center for Commercial Agriculture. <https://ag.purdue.edu/commercialag/home/paer-publication/2024-08-paer-purdue-farmland-values-and-cash-rents-survey-results/.>

Ogle, T., & DeBoer, L. (2018, August 1). Farmland assessment for property taxes in the coming decade. Purdue University Center for Commercial Agriculture. <https://ag.purdue.edu/commercialag/home/paer-article/farmland-assessment-for-property-taxes-in-the-coming-decade.>
